% Set the author and title of the compiled pdf
\hypersetup{
	pdftitle = {\Title},
	pdfauthor = {\Author}
}

\section{Course intro}

A smartphone is a mobile phone running a mobile operating system, with advanced
capabilities with regard to computing power and connectivity. They are much more
advanced than traditional feature phones, and have lots of features, including
cameras, multimedia functionality, GPS, touch screens etc.

There are three main mobile operating systems in use today:

\begin{description}

  \item \textbf{Android}\\
  	Founded in 2003 by Andy Rubin and backed by Google. It's mostly free and
  	open source, and holds a very strong position in the market.
  \item \textbf{iOS}\\
  	Introduced in 2007 by Apple, this is a closed source operating system. The
  	first iOS phones were very groundbreaking in terms of their technology, and
  	were the first to feature touch screens, which are now ubiqutous in the
  	market.
  \item \textbf{Windows Phone}\\
    Version seven was released in 2010, previous versions were terrible (imho).
  \item \textbf{Blackberry}\\
   	%TODO: Write stuff
\end{description}

\subsection{What's in a smartphone?}

Battery, memory, touch screen, radio, mic, etc

\subsection{Breakdown of the course}

%TODO: Only fill this in if you're bored


% Lecture two
\section{Signals in mobile systems}

Signals such as speech and music arrive at the device as physical, analogue quantities that vary in a continous manner over time.

% Worksop 1
% Narrow band soeech coding for mobile phones