\section*{Introduction}

Now that the mobile telephone has evolved into a powerful computer, the mobile
dimension of computing is a vital part of Computer Science. This unit will give
insights into many issues of mobile systems, including wireless communication
networks, the processing of speech, music and other real-time signals, the
control of bit-errors and maximising battery life. The techniques and software
which underlie commonplace applications of mobile computing systems, including
smart-phones, tablets, laptop computers, MP3 players and GPS satellite
navigation, will be addressed.

\section*{Aims}

Computing is becoming increasingly mobile. This unit will give insights into the
issues of mobile systems, covering mobile communications, real-time signals such
as speech, video and music, codecs, and maximising battery life.

\begin{mymulticols}
  \begin{itemize}
    \item Commonplace examples of mobile computing systems: - mobile phones; - MP3 players; - laptop computers; - PDAs; - GPS satellite navigation.
    \item Real-time signals
    \item Analogue and digital signals; - time and frequency domain representations; - sampling, aliasing, quantization; - companding; - real-time computation.
    \item Coding, decoding and compression
    \item GSM speech coding; - MP3 music, JPEG image and MPEG video coding \& decoding; - error correcting codes; - communications coding schemes.
    \item Mobile communication
    \item Transmitting real-time information over wireless networks; - principles of cellular and ad-hoc networks; - Coding of multimedia signals - to increase the capacity of radio channels; - to minimise the effect of transmission errors.
    \item Maximising battery life
    \item May be addressed at many levels including: - chip design; - signal coding and processing; - medium access control; - transmit power control.
  \end{itemize}
\end{mymulticols}

\section*{Contributing \& Attribution}

These notes are open-sourced on Github at
\url{https://github.com/Todd-Davies/second-year-notes}. Please feel free to
submit a pull request if you want to make any changes, or maybe open an issue
if you find an error! Feedback is very welcome; you can email me at
\href{mailto:todd434@gmail.com}{todd434@gmail.com}.

These notes are also based off the \texttt{COMP28512} course notes by Dr. Barry
Cheetham and Dr. Steve Furber.
